---
title: "Take-home Exercise 2"
author: "Wang Ruipeng"
date: "Jan 25, 2023"
---

# 1.0 Overview

This take-home exercise is done based on a take-home exercise 1 submission prepared by a classmate. The peer submission will be critiqued in terms of clarity and aesthetics, and the original design will be remade using the data visualization principles and best practice learnt in Lesson 1 and 2.

# 2.0 Data Preparation

In this step, we will load the libraries and dataset

```{r}
# install and load libraries 
packages = c('tidyverse', 'knitr', 'ggdist', 'scales', 'grid', 'gridExtra',
             'formattable', 'patchwork', 'plotly')
for(p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p, character.only = T)
}
```

```{r}
df <- read_csv('data/respopagesextod2022.csv')
```

```{r}
library(rmarkdown)
paged_table(df)
```

# 3.0 Visualization Remake Using R

### 3.1 Exploration using `ggplot()`

```{r}
ggplot(df, aes (x = AG, y = Pop, fill = Sex)) +
  geom_bar(stat = "identity") +
  coord_flip()
```

### 3.2 Data Wrangling

Transforming the data to one gender.

```{r}
Popdata_males <- df %>%
  filter(`Sex` ==  "Males") %>%
  mutate (Pop = -Pop)
glimpse(Popdata_males)
```

```{r}
Popdata_females <-df %>%
  filter(`Sex` ==  "Females") 
Popdata_T <- rbind(Popdata_males,Popdata_females)
paged_table(Popdata_T)
```

### 3.3 Initial Pyramid Plot 

```{r}
ggplot(Popdata_T, aes (x = AG, y = Pop, fill = Sex)) +
  geom_bar(stat = "identity") +
  coord_flip()
```

### 3.4 Planning Areas

```{r}
print(unique(Popdata_T$PA))
```

### 3.5 Trellis display

```{r}
xbrks <- seq(-50000, 50000, 10000)
xlabls <- paste0(as.character(c(seq(50, 0, -10), seq(10, 50, 10))))

p <- ggplot(subset(Popdata_T, PA %in% c("Ang Mo Kio", "Sengkang", "Tampines", "Jurong East", "Bishan", "Woodlands","Yishun","Punggol","Outram")), 
      aes (x = AG, y = Pop, fill = Sex)) +
      geom_bar(stat = "identity", width = .9 ) +
      scale_y_continuous(breaks = xbrks, labels = xlabls, name = "Population Size ('000)") +
      xlab("Age Groups") +
      coord_flip() +
      theme_bw() 
```

```{r fig.height = 10, fig.width = 7}
xbrks1 <- seq(-50000, 50000, 10000)
xlabls1 <- paste0(as.character(c(seq(50, 0, -10), seq(10, 50, 10))))
p1 <- p +  facet_wrap (~ PA, nrow = 3, ncol = 3) + 
            xlab("Age Groups") +
            ggtitle("Age-Sex Pyramid for Selected 9 Planning Areas by Population")+
            scale_y_continuous(breaks = xbrks1, labels = xlabls1, name = "Population Size ('000)")

ggplotly(p1)
```

Source: [*Singapore Residents by Planning Area / Subzone, Age Group, Sex and Type of Dwelling, June 2022*](https://www.singstat.gov.sg/)
